Flatiron cleaning

### 2021-12-16 Rmd/rdi analysis.Rmd
check and clean code rmd

@@

- Select episodedatasource == "Administrations"
```{r drugs episode cleanup}
drugs <- drugs %>% 
  filter(episodedatasource == "Administrations" #& 
           # ismaintenancetherapy == "FALSE" # keep maintenance for a correct recoding of line number
         ) %>%
  # bind with clinical
  inner_join(., clinical_data, 
          by = "patientid") %>% 
  arrange(patientid, episodedate)

drugs %>% distinct(patientid) %>% nrow()
```

- Combine multiple administrations of the same drug at the same date
Flatiron says:
There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.  
For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.  
Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend.   
Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
```{r combine amount same drug one day}
drugs1 <- drugs %>% 
  # 3. Combined Multiple Administrations----
  # combined multiple amount of drug given the same day FCF86329BB40C F4E6EE1910940.May be 205 patients row like that
# There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.
# For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.
# Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend. 
# Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
  group_by(patientid, linename, linenumber, linestartdate, lineenddate, episodedate, 
           drugname, units, ismaintenancetherapy) %>%
  summarise_at(vars(route, amount), paste, collapse = ";") %>% # don't do sum ti keep the multiple administration in df
  mutate(route = case_when(
    str_detect(route, ";")            ~ "Intravenous",
    TRUE                              ~ route
  )) %>% 
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  ungroup() %>% 
  mutate(across(starts_with("amount"), ~ na_if(., "NA"))) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  mutate(across(starts_with("amount"), ~ as.numeric(.))) %>% 
  mutate(amount = rowSums(select(.,amount_1:amount_4), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, and NA they are all good, but 
  # ug need conversion
  # the AUC are mostly correct (real mg) but are actual AUC when < at 10 so remove
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  )) %>%

  mutate(drugname_type = case_when(
    str_detect(drugname, "platin")       ~ "platin",
    str_detect(drugname, "taxel")        ~ "taxel"
  ))
```

- Recode linenumber depending on surgery date (Categorize Neo and Adjuavnt)
```{r line coding}
therapy_line <- drugs1 %>% 
  # bind again with clinical - easier that way
  inner_join(., clinical_data, 
          by = "patientid") %>% 
  arrange(patientid, episodedate) %>% 
  # 4. Define Adjuvant, Neoadjuvant, chemo or surgery----
  mutate(chemotherapy_type = case_when(
    # ismaintenancetherapy == "TRUE"    ~ "maintenance",
    episodedate <= surgerydate        ~ "Neoadjuvant", # F63ABD6AEB12C, F4F7B1C5DF24F intraperitoneal Carboplatin
    episodedate > surgerydate         ~ "Adjuvant",
    is.na(surgerydate) &
      !is.na(episodedate)             ~ "Chemotherapy only",
    !is.na(surgerydate) &
      is.na(episodedate)              ~ "Surgery only"
    )) %>% 
  mutate(chemotherapy_type = 
           factor(chemotherapy_type, levels = 
                    c("Neoadjuvant", "Adjuvant", "Chemotherapy only", "Surgery only"))) %>%
  group_by(patientid) %>% 
  mutate(had_neo = ifelse(str_detect(chemotherapy_type, "Neo"), "Yes", NA_character_)) %>% 
  fill(had_neo, .direction = "updown") %>% 
  mutate(had_adj = ifelse(str_detect(chemotherapy_type, "Adj"), "Yes", NA_character_)) %>% 
  fill(had_adj, .direction = "updown") %>% 
  mutate(treatment_sequence = case_when(
    had_neo == "Yes" &
      had_adj == "Yes"                           ~ "neo + surg + adj",
    had_neo == "Yes" &
      is.na(had_adj)                             ~ "neo + surg",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "surg + adj",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only"
  )) %>% 
  mutate(therapy = case_when(
    chemotherapy_type == "Neoadjuvant"           ~ "Upfront Chemo",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "Upfront Surgery"
  )) %>% 
  fill(therapy, .direction = "downup") %>%

  # 5.Redefine line number based on the original linenumber variable and the surgery date----
  mutate(linenumber_n = dense_rank(interaction(linenumber, chemotherapy_type))) %>%  
  group_by(patientid, linenumber_n, ismaintenancetherapy) %>% 
  mutate(maint_rank = row_number(ismaintenancetherapy),
         maint_rank = ifelse(maint_rank == 1, 1, 0)) %>%

  group_by(patientid) %>% 
  mutate(linenumber_new = cumsum(maint_rank)) %>%  
  # F0000AD999ABF, F1C68CFAC2AF3, FBCF69031DFF8, F0040EA299CF0, FAACDC7205803,  FA019D2071321 no has big jump F95D860690A93 still weird F0F1324B863A8
  select(patientid, linename, linenumber, linenumber_new, drugname, episodedate, surgerydate, chemotherapy_type, everything()) %>% 
  ungroup() %>%

  group_by(patientid, chemotherapy_type) %>% 
  mutate(linenumber_adj = dense_rank(linenumber_new)) %>% 
  ungroup() %>% 
  # 6.Add a 30 day rule to switch linenumber----
  # group_by(patientid, linenumber_new) %>% 
  # arrange(episodedate) %>% 
  # mutate(episode_interval = episodedate - lag(episodedate), episode_interval = ifelse(is.na(episode_interval), 0, episode_interval)) %>%  # FDB58481AFFBB 48 days
  # mutate(jump_line = ifelse(episode_interval > 31, 1, 0)) %>% # FA019D2071321 >31
  # ungroup() %>% 
  # group_by(patientid) %>% 
  # mutate(jump_line = cumsum(jump_line)) %>% # F002DD2953889 fixed even when increase multiple time 
  # mutate(linenumber_new = linenumber_new + jump_line) %>% 
  #F002DD2953889 need to do 37 and 42 F0040EA299CF0, 35 F005602DA4DF0 => do at least 37

  # 7.Calculate cycle for each line----
  mutate(cycle_drugname = str_remove(drugname, " protein-bound")) %>%  # F003F2BEEDB37 need to change drugname pacli = pacli bp
  group_by(patientid, linenumber_new, cycle_drugname) %>% # Wrong line name F001443D8B85C ~~~~~~~~~~~~~~~~~~~~~~~~~~ Fix later QUESTION Dr Chern is ok
  mutate(#drugname_count_perline = n(),
         drugname_count_perline = case_when(
           str_detect(drugname, "carboplatin|paclitaxel")     ~ n()
         )
           ) %>%  # F0000AD999ABF
  ungroup() %>% 
  mutate(across(c("linenumber", "linenumber_new", "linenumber_adj"), as.factor)) %>% 
  group_by(patientid, linenumber_new) %>% 
  mutate(linestartdate_new = min(episodedate)) %>% 
  mutate(lineenddate_new = max(episodedate)) %>% 
  ungroup()
```

```{r cycle coding}
# Code cycle number
therapy_cycle <- therapy_line %>% 
  arrange(patientid, episodedate, drugname) %>% 
  group_by(patientid, linenumber_new) %>%
  # Need the number of time of the "major" drug is seen in the line
  mutate(cycle_count_perline = min(drugname_count_perline, na.rm = TRUE)) %>%
  mutate(cycle_major_drugname = case_when(
    cycle_count_perline == drugname_count_perline &
      (cycle_drugname == "carboplatin" |
      cycle_drugname == "paclitaxel")                      ~ cycle_drugname,
    TRUE                                                   ~ NA_character_
  )) %>% 
  # select(-drugname_count_perline) %>% 
  group_by(patientid, linenumber_new, cycle_major_drugname) %>%
  mutate(cycle_increment = row_number(cycle_major_drugname)) %>% # Look at F002DD2953889 had a long time without one drug------------------
  # F01D908A013B7 has more pacli at the end but in the same line

  # + What to do when it is a cycle change F003F2BEEDB37 ,

  # group_by(patientid, linenumber_new, cycle_inc) %>%
  # mutate(cycle_count_1 = row_number(cycle_inc)) %>% 
  # mutate(cycle_c = ifelse(cycle_count_1 == 1, cycle_inc, NA_real_)) %>% 
  arrange(patientid, episodedate, cycle_increment) %>% 
  group_by(patientid, linenumber_new) %>%
  # mutate(cycle_increment = cycle_inc) %>% 
  fill(cycle_increment, .direction = "downup") %>%  # make sure I can do up for have one new drug compare to the linename FBB5186A15CD9, FE593EDF5586A
  group_by(patientid, chemotherapy_type, cycle_increment) %>% 
  mutate(is_dose_dense = row_number(cycle_increment),
         is_dose_dense = ifelse(is_dose_dense == 1, 1, NA_real_)) %>% 
  group_by(patientid) %>% 
  mutate(total_cycle_increment = #case_when(
           # ismaintenancetherapy == "FALSE"            ~ 
           row_number(is_dose_dense)
         # )
  ) %>% 
  fill(total_cycle_increment, .direction = "downup") %>% 
  # cycle date
  group_by(patientid, linenumber_new, cycle_increment) %>% 
  mutate(cycle_start_date = min(episodedate)) %>%

  # Look at dose dense
  group_by(patientid, linenumber_new, cycle_increment, cycle_drugname) %>%  # weird dose dense F064D9EB733B3 F050B815FC126
  mutate(dose_dense = n()) %>% 
  mutate(dose_dense = max(dose_dense)) %>%  # too much cycle
  ungroup()

@@

```{r rule between diagnosis and treatment}
Treatment <- Treatment1 %>% 
  filter(route == "Intravenous") %>% 
  # 1. 60 days rule between diagnosis and treatment # Loss 1
  group_by(patientid) %>% 
  mutate(first_drug_date = first(episodedate)) %>% 
  mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "days")) %>% 
  mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "days")) %>% 
  mutate(days_at_first_treatment = case_when(
    therapy == "Upfront Chemo" |
      therapy == "Chemotherapy only"             ~ interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "days"),
    therapy == "Upfront Surgery" |
      therapy == "Surgery Only"                  ~ interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "days")
  )) %>% 
  ungroup() %>% 
  mutate(thirty_days_exclusion = interval(start = diagnosisdate, end = dateofdeath)/
           duration(n=1, units = "days")
  ) %>% 
  group_by(patientid) %>% 
  mutate(had_carbo_pacli = case_when(
    drugname == "paclitaxel" |
      drugname == "carboplatin"             ~ "Yes"
  )) %>% 
  fill(had_carbo_pacli, .direction = "updown") %>% 
  ungroup() %>% 
  mutate(interval_surgery_to_adj_start = case_when(
      chemotherapy_type == "Adjuvant" &
        linenumber_adj == 1       
      ~ interval(start = surgerydate, end = linestartdate_new) / duration(n = 1, units = "months")
    ),
  ) %>% 
  mutate(ex = case_when(
    interval_surgery_to_adj_start > 3                  ~ "exclude"
  )) %>% 
  group_by(patientid) %>% 
  fill(ex, .direction = "updown") %>% 
  ungroup() %>%

  group_by(patientid) %>% 
  mutate(neo_more_more_than_1line = case_when(
    chemotherapy_type == "Neoadjuvant" &
           linenumber_new != 1                 ~ "exclude"
  )) %>% 
  fill(neo_more_more_than_1line, .direction = "updown") %>% 
  ungroup() %>% 
  mutate(whole_frontline_is_maintenance = case_when(
    (linenumber_new == 1 &
    ismaintenancetherapy == "TRUE") |
    (linenumber_new == 2 &
      therapy == "Upfront Chemo" &
    ismaintenancetherapy == "TRUE")            ~ "exclude"
  )) %>% 
  group_by(patientid) %>% 
  fill(whole_frontline_is_maintenance, .direction = "updown") %>% 
  ungroup() %>% 
  mutate(too_many_cycle = case_when(
    (linenumber_new == 1 &
       therapy == "Upfront Chemo" &
       total_cycle_increment > 6) |

@@

      (linenumber_new == 1 &
         therapy == "Upfront Surgery" &
         total_cycle_increment > 9)             ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(too_many_cycle, .direction = "updown") %>%
  ungroup() %>%

  mutate(wrong_pacli_dosedense = case_when(
    drugname == "paclitaxel" &
      is.na(too_many_cycle) &
      (dose_dense == 2 |
      dose_dense > 3) &

      ((linenumber_new == 1 &
       therapy == "Upfront Chemo") |

      (linenumber_new == 2 &
         therapy == "Upfront Chemo") |

      (linenumber_new == 1 &
         therapy == "Upfront Surgery"))
    ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(wrong_pacli_dosedense, .direction = "updown") %>% 
  ungroup() %>% 
  mutate(switch_drugs = case_when(
    str_detect(linename, "Pacli") & 
      drugname == "docetaxel" & 
      linenumber_adj == 1                      ~ "exclude",
    str_detect(linename, "Car") & 
      drugname == "cisplatin" & 
      linenumber_adj == 1                      ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(switch_drugs, .direction = "updown") %>% 
  ungroup() %>%

  mutate(line_keep = case_when(# F002B429BE6C6
    (linenumber_adj == 1 &
           (linename == "Carboplatin,Paclitaxel" |
           linename == "Bevacizumab,Carboplatin,Paclitaxel" |
           linename == "Carboplatin" |
           linename == "Paclitaxel" |
           linename == "Bevacizumab,Paclitaxel" |
           linename == "Bevacizumab,Carboplatin"))            ~ NA_character_,
    linenumber_adj == 1                                       ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(line_keep, .direction = "updown") %>% 
  ungroup() %>% 
  mutate(flatiron_rules = case_when(
    days_at_first_treatment > 90                          ~ "exclude",
    thirty_days_exclusion <= 30                  ~ "exclude",
    exclude == "1"                               ~ "exclude"
  )) %>%
  group_by(patientid) %>% 
  fill(flatiron_rules, .direction = "downup") %>% 
  ungroup()

@@

missing_data <- missing_data %>%
  filter(is.na(ex))

@@

## 4. step 4: Include patients with treatments following the standart of clinical care

- exclude patients with more than 1 line as Neo
- exclude patients with maintenance within their frontline
- exclude patients with more than 6 cycles in Neo, more than 9 cycles **total** in upfront chemo or more than 9 cycles in upfront surgery
```{r}
standart_care <- missing_data %>% 
  # Remove patients with more than 1 line as Neo
  filter(is.na(neo_more_more_than_1line))

standart_care <- standart_care %>% 
  filter(is.na(whole_frontline_is_maintenance))

standart_care <- standart_care %>% 
  filter(is.na(too_many_cycle))
```
**We have `r standart_care %>% distinct(patientid) %>% nrow()` patients.**

- Exclude patients that received wrong dose dense of 1 carbo + 2 paclitaxels
```{r}
Frontline <- standart_care %>% 
  filter(is.na(wrong_pacli_dosedense))

# standart_care %>% 
#   mutate(n_dose_dense = factor(n_dose_dense)) %>% 
#   select(n_dose_dense, therapy) %>% 
#   tbl_summary(by = therapy)

>>>>

- Select episodedatasource == "Administrations"

@@

```{r drugs episode cleanup}
drugs <- drugs %>% 
  # Select only records for administration
  filter(episodedatasource == "Administrations") %>%
  # bind with clinical
  inner_join(., clinical_data, 
          by = "patientid") %>% 
  arrange(patientid, episodedate)
```
But when merged both data, we have `r drugs %>% distinct(patientid) %>% nrow()` patients with clinical informations who received drugs.

- Combine multiple administrations of the same drug at the same date
Flatiron says:
There are a small number of patients with multiple records in the MedicationAdministration table for the same drug on the same day. We reviewed several instances where this occurs in the original EMR records and found, in each case, legitimate clinical reasons substantiating the multiple administrations per day. Although these instances were legitimate medication administrations, Flatiron Health data are generated from real-world clinical practice and are subject to miscoding and errors by the clinical oncology team, as are all EMR data. Below are three patient vignettes taken from our review of EMR notes, illustrating the variety of reasons why this may occur.  
For clinical reasons, a patient was given higher than normal dose which was listed as two administrations of the normal dose.  
Three administrations of the same drug were listed on a Friday. The EMR note stated that one dose was administered in the office and the patient was sent home with the two remaining doses for the weekend.   
Patient had an allergic reaction in the past and was administered multiple smaller doses in the same day as part of a desensitization protocol.
```{r combine amount same drug one day}
combined_drugs <- drugs %>% 
  # Combined Multiple Administrations----
  # combined multiple amount of drug given the same day FCF86329BB40C F4E6EE1910940.May be 205 patients row like that
  group_by(patientid, linename, linenumber, linestartdate, lineenddate, episodedate,
           drugname, units, ismaintenancetherapy) %>%
  summarise_at(vars(route, amount), paste, collapse = ";") %>% # keep them in the data then do sum
  mutate(route = case_when(
    str_detect(route, ";")            ~ "Intravenous", # I checked that all the combined administration are IV
    TRUE                              ~ route
  )) %>% 
  separate(col= amount, paste("amount_", 1:10, sep=""), sep = ";", extra = "warn",
           fill = "right") %>%
  ungroup() %>% 
  mutate(across(starts_with("amount"), ~ na_if(., "NA"))) %>% 
  purrr::keep(~!all(is.na(.))) %>%
  mutate(across(starts_with("amount"), ~ as.numeric(.))) %>% 
  mutate(amount = rowSums(select(.,starts_with("amount_")), na.rm = TRUE)) %>%
  # Found ug, ml, Ea, and NA they are all good, but 
  # ug need conversion
  # the AUC are mostly correct (real mg) but are actual AUC when < at 10 so remove
  mutate(amount =  case_when(
    units == "ug"               ~ amount / 10, # I checked all of them
    units == "AUC" &
      amount < 10               ~ NA_real_,
    TRUE                        ~ amount
  ))# %>%

  # mutate(drugname_type = case_when(
  #   str_detect(drugname, "platin")       ~ "platin",
  #   str_detect(drugname, "taxel")        ~ "taxel"
  # ))
```

- Recode linenumber depending on surgery date (Categorize Neo and Adjuavnt)
```{r line coding}
therapy_line <- combined_drugs %>% 
  # bind again with clinical - was lost while summarizing
  inner_join(., clinical_data, 
          by = "patientid") %>% 
  arrange(patientid, episodedate) %>% 
  # Define Adjuvant, Neoadjuvant, chemo or surgery
  mutate(chemotherapy_type = case_when(
    # ismaintenancetherapy == "TRUE"    ~ "maintenance",
    episodedate <= surgerydate        ~ "Neoadjuvant", # F63ABD6AEB12C, F4F7B1C5DF24F intraperitoneal Carboplatin
    episodedate > surgerydate         ~ "Adjuvant",
    is.na(surgerydate) &
      !is.na(episodedate)             ~ "Chemotherapy only",
    !is.na(surgerydate) &
      is.na(episodedate)              ~ "Surgery only"
    )) %>% 
  mutate(chemotherapy_type = 
           factor(chemotherapy_type, levels = 
                    c("Neoadjuvant", "Adjuvant", "Chemotherapy only", "Surgery only"))) %>%
  group_by(patientid) %>% 
  mutate(had_neo = ifelse(str_detect(chemotherapy_type, "Neo"), "Yes", NA_character_)) %>% 
  mutate(had_adj = ifelse(str_detect(chemotherapy_type, "Adj"), "Yes", NA_character_)) %>% 
  fill(had_neo, had_adj, .direction = "updown") %>% 
  mutate(treatment_sequence = case_when(
    had_neo == "Yes" &
      had_adj == "Yes"                           ~ "neo + surg + adj",
    had_neo == "Yes" &
      is.na(had_adj)                             ~ "neo + surg",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "surg + adj",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only"
  )) %>% 
  mutate(therapy = case_when(
    chemotherapy_type == "Neoadjuvant"           ~ "Upfront Chemo",
    chemotherapy_type == "Chemotherapy only"     ~ "Chemotherapy only",
    chemotherapy_type == "Surgery only"          ~ "Surgery only",
    is.na(had_neo) &
      had_adj == "Yes"                           ~ "Upfront Surgery"
  )) %>% 
  fill(therapy, .direction = "downup") %>%

  # Redefine line number based on the original linenumber variable and the surgery date
  mutate(linenumber_n = dense_rank(interaction(linenumber, chemotherapy_type))) %>%  
  # Redefine line number based on the new linenumber variable and if a switch 
  # to maintenance is detected within the line
  group_by(patientid, linenumber_n, ismaintenancetherapy) %>% 
  mutate(maint_rank = row_number(ismaintenancetherapy),
         maint_rank = ifelse(maint_rank == 1, 1, 0)) %>% 
  group_by(patientid) %>% 
  mutate(linenumber_new = cumsum(maint_rank)) %>%  
  # F0000AD999ABF, F1C68CFAC2AF3, FBCF69031DFF8, F0040EA299CF0, FAACDC7205803,  FA019D2071321 no has big jump F95D860690A93 F0F1324B863A8
  select(patientid, linename, linenumber, linenumber_new, drugname, 
         episodedate, surgerydate, chemotherapy_type, everything()) %>% 
  ungroup() %>% 
  # Define line number restarting to 1 for each chemotherapy_type
  group_by(patientid, chemotherapy_type) %>% 
  mutate(linenumber_adj = dense_rank(linenumber_new)) %>% 
  ungroup() %>% 
  mutate(across(c("linenumber", "linenumber_new", "linenumber_adj"), as.factor)) %>% 
  # Redefine line dates
  group_by(patientid, linenumber_new) %>% 
  mutate(linestartdate = min(episodedate)) %>% 
  mutate(lineenddate = max(episodedate)) %>% 
  ungroup() %>% 
  select(-c(linenumber_n, maint_rank))

  # 6.Add a 30 day rule to switch linenumber----
  # group_by(patientid, linenumber_new) %>% 
  # arrange(episodedate) %>% 
  # mutate(episode_interval = episodedate - lag(episodedate), episode_interval = ifelse(is.na(episode_interval), 0, episode_interval)) %>%  # FDB58481AFFBB 48 days
  # mutate(jump_line = ifelse(episode_interval > 31, 1, 0)) %>% # FA019D2071321 >31
  # ungroup() %>% 
  # group_by(patientid) %>% 
  # mutate(jump_line = cumsum(jump_line)) %>% # F002DD2953889 fixed even when increase multiple time 
  # mutate(linenumber_new = linenumber_new + jump_line) %>% 
  #F002DD2953889 need to do 37 and 42 F0040EA299CF0, 35 F005602DA4DF0 => do at least 37

```

```{r cycle coding}
therapy_cycle <- therapy_line %>% 
  # Define cycle for each line
  # F003F2BEEDB37 need to change drugname pacli = pacli bp
  mutate(cycle_drugname = str_remove(drugname, " protein-bound")) %>%  
  # Need the number of time of the "major" drug is seen in the line
  group_by(patientid, linenumber_new, cycle_drugname) %>%
  mutate(drugname_count_perline = case_when(
           str_detect(drugname, "carboplatin|paclitaxel")     ~ n()
         )) %>%
  ungroup() %>% 
  arrange(patientid, episodedate, drugname) %>% 
  group_by(patientid, linenumber_new) %>%
  # Take the major drugname as a mark of new cycle start
  mutate(cycle_count_perline = min(drugname_count_perline, na.rm = TRUE)) %>%
  mutate(cycle_major_drugname = case_when(
    cycle_count_perline == drugname_count_perline &
      (cycle_drugname == "carboplatin" |
      cycle_drugname == "paclitaxel")                      ~ cycle_drugname,
    TRUE                                                   ~ NA_character_
  )) %>% 
  # select(-drugname_count_perline) %>% 
  # Code cycle number based on when major drug name appear again
  group_by(patientid, linenumber_new, drugname, cycle_major_drugname) %>%
  mutate(cycle_increment = row_number(cycle_major_drugname)) %>% # Look at F002DD2953889 inversed dose dense------------------
  # F01D908A013B7 weird dose dense

  # F003F2BEEDB37 change of dose dense 50% ok? ---------------------------------------------------------------

  arrange(patientid, episodedate, cycle_increment) %>% 
  group_by(patientid, linenumber_new) %>%
  fill(cycle_increment, .direction = "downup") %>%  # make sure I can do up for have one new drug compare to the linename FBB5186A15CD9 ok, FE593EDF5586A has a switched drugs will be removed
  group_by(patientid, chemotherapy_type, cycle_increment) %>% 
  # Code is_dose_dense to be 1 for the first drug in the cycle
  mutate(is_dose_dense = row_number(cycle_increment),
         is_dose_dense = ifelse(is_dose_dense == 1, 1, NA_real_)) %>% 
  group_by(patientid) %>% 
  mutate(total_cycle_increment = row_number(is_dose_dense)
         ) %>% 
  group_by(patientid, linenumber_adj) %>% 
  fill(total_cycle_increment, .direction = "downup") %>% 
  # cycle date
  group_by(patientid, linenumber_new, cycle_increment) %>% 
  mutate(cycle_start_date = min(episodedate)) %>%

  # Look at dose dense
  # Code how many time a drug comes up in a cycle
  group_by(patientid, linenumber_new, cycle_increment, cycle_drugname) %>%  # weird dose dense F064D9EB733B3 F050B815FC126
  mutate(dose_dense = n()) %>% 
  mutate(dose_dense = max(dose_dense)) %>%  # too much cycle
  ungroup()

@@

```{r rule between diagnosis and treatment}
Treatment <- Treatment1 %>% 
  filter(route == "Intravenous") %>% 
  # step 1: Flatiron recommendation cleanup
  # 90 days rule between diagnosis and treatment
  group_by(patientid) %>% 
  mutate(first_drug_date = first(episodedate)) %>% 
  ungroup() %>% 
  mutate(days_at_first_drug = interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "days")) %>% 
  mutate(days_at_surg = interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "days")) %>% 
  mutate(days_at_first_treatment = case_when(
    therapy == "Upfront Chemo" |
      therapy == "Chemotherapy only"             ~ interval(start = diagnosisdate, end = first_drug_date)/
           duration(n=1, units = "days"),
    therapy == "Upfront Surgery" |
      therapy == "Surgery Only"                  ~ interval(start = diagnosisdate, end = surgerydate)/
           duration(n=1, units = "days")
  )) %>% 
  # 30 days rule excluding patient who didn’t survive for at least 30 days after diagnosis
  mutate(thirty_days_exclusion = interval(start = diagnosisdate, end = dateofdeath)/
           duration(n=1, units = "days")
  ) %>% 
  mutate(flatiron_rules = case_when(
    days_at_first_treatment > 90                 ~ "exclude",
    thirty_days_exclusion <= 30                  ~ "exclude",
    exclude == "1"                               ~ "exclude"
  )) %>%
  group_by(patientid) %>% 
  fill(flatiron_rules, .direction = "downup") %>% 
  ungroup() %>%

  # step 2: Filter patients with clinical characteristics fitting to the study criteria
  # include only patients who received platin and/or taxane
  mutate(had_carbo_pacli = case_when(
    drugname == "paclitaxel" |
      drugname == "carboplatin"             ~ "Yes"
  )) %>% 
  group_by(patientid) %>% 
  fill(had_carbo_pacli, .direction = "updown") %>% 
  ungroup() %>% 
  # include patients with single or dual P/T +/- Bev as frontline treatment
  mutate(line_keep = case_when(# F002B429BE6C6
    (linenumber_adj == 1 &
           (linename == "Carboplatin,Paclitaxel" |
           linename == "Bevacizumab,Carboplatin,Paclitaxel" |
           linename == "Carboplatin" |
           linename == "Paclitaxel" |
           linename == "Bevacizumab,Paclitaxel" |
           linename == "Bevacizumab,Carboplatin"))            ~ NA_character_,
    linenumber_adj == 1                                       ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(line_keep, .direction = "updown") %>% 
  ungroup() %>%

  # step 3: Filter out patients with missing data
  # remove patients with long gap between surgery and adjuvant start (3 months)
  mutate(interval_surgery_to_adj_start = case_when(
      chemotherapy_type == "Adjuvant" &
        linenumber_adj == 1       
      ~ interval(start = surgerydate, end = linestartdate) / duration(n = 1, units = "months")
    )) %>% 
  mutate(gap_surg_drugs = case_when(
    interval_surgery_to_adj_start > 3                  ~ "exclude"
  )) %>% 
  group_by(patientid) %>% 
  fill(gap_surg_drugs, .direction = "updown") %>% 
  ungroup() %>%

  # step 4: Include patients with treatments following the standard of clinical care
  # exclude patients with more than 1 line as Neo
  mutate(neo_more_more_than_1line = case_when(
    chemotherapy_type == "Neoadjuvant" &
           linenumber_new != 1                 ~ "exclude"
  )) %>% 
  group_by(patientid) %>% 
  fill(neo_more_more_than_1line, .direction = "updown") %>% 
  ungroup() %>% 
  # exclude patients with maintenance within their frontline
  mutate(frontline_has_maintenance = case_when(
    linenumber_adj == 1 &
    ismaintenancetherapy == "TRUE"             ~ "exclude"
  )) %>% 
  group_by(patientid) %>% 
  fill(frontline_has_maintenance, .direction = "updown") %>% 
  ungroup() %>% 
  # exclude patients with more than 6 cycles in Neo, more than 9 cycles total 
  # in upfront chemo or more than 9 cycles in upfront surgery
  mutate(too_many_cycle = case_when(
    (linenumber_new == 1 &
       therapy == "Upfront Chemo" &
       total_cycle_increment > 6) |

@@

      (linenumber_new == 1 &
         therapy == "Upfront Surgery" &
         total_cycle_increment > 9)             ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(too_many_cycle, .direction = "updown") %>%
  ungroup() %>% 
  # Exclude patients that received wrong dose dense of 1 carbo + 2 paclitaxels
  # Removes also the patients who has only carbo or only pacli
  # dose_dense will need to be changed if we change and include these patients
  arrange(patientid, episodedate, drugname) %>% 
  mutate(wrong_pacli_dosedense = case_when(
    drugname == "paclitaxel" &
      is.na(too_many_cycle) &
      (dose_dense == 2 |
      dose_dense > 3) &
      linenumber_adj == 1
      # ((linenumber_new == 1 &
      #  therapy == "Upfront Chemo") |
      # (linenumber_new == 2 &
      #    therapy == "Upfront Chemo") |
      # (linenumber_new == 1 &
      #    therapy == "Upfront Surgery"))
    ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(wrong_pacli_dosedense, .direction = "updown") %>% 
  ungroup() %>% 
  # Exclude patients who switch drug within a line
  mutate(switch_drugs = case_when(
    str_detect(linename, "Pacli") & 
      drugname == "docetaxel" & 
      linenumber_adj == 1                      ~ "exclude",
    str_detect(linename, "Car") & 
      drugname == "cisplatin" & 
      linenumber_adj == 1                      ~ "exclude"
  )) %>% 
  group_by(patientid) %>%
  fill(switch_drugs, .direction = "updown") %>% 
  ungroup()

@@

missing_data <- missing_data %>%
  filter(is.na(gap_surg_drugs))

@@

## 4. step 4: Include patients with treatments following the standard of clinical care

- exclude patients with more than 1 line as Neo
- exclude patients with maintenance within their frontline
- exclude patients with more than 6 cycles in Neo, more than 9 cycles **total** in upfront chemo or more than 9 cycles in upfront surgery
```{r}
standard_care <- missing_data %>% 
  # Remove patients with more than 1 line as Neo
  filter(is.na(neo_more_more_than_1line))

standard_care <- standard_care %>% 
  filter(is.na(frontline_has_maintenance))

standard_care <- standard_care %>% 
  filter(is.na(too_many_cycle))
```
**We have `r standard_care %>% distinct(patientid) %>% nrow()` patients.**

- Exclude patients that received wrong dose dense of 1 carbo + 2 paclitaxels
```{r}
Frontline <- standard_care %>% 
  filter(is.na(wrong_pacli_dosedense))

# standard_care %>% 
#   mutate(n_dose_dense = factor(n_dose_dense)) %>% 
#   select(n_dose_dense, therapy) %>% 
#   tbl_summary(by = therapy)
